---
title: 'TidyTuesday 2020 week 16: Best Rap Artists'
subtitle: https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-04-14/readme.md
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r setup}
pckgs <- c("tidyverse","lubridate", "here")

invisible(lapply(pckgs, library, character.only = TRUE))
```

Data Input:

```{r input, results='hide'}
polls <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/polls.csv')
rankings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/rankings.csv')
```

Lets look at the data:

```{r}
glimpse(rankings)
```

Any missing values?
```{r}
rankings %>% 
  map_lgl(~any(is.na(.))) %>% 
  any()
  
```


# Rankings

```{r}
summary(rankings)
```

```{r}
rankings %>% n_distinct(title)
```

311 songs - each ranked `n` times with the total amount of points equal to `points` .

```{r}
rankings <- rankings %>% 
  mutate(decade = (year %/% 10)*10)
```

```{r}
rankings %>% 
  count(decade)
```

Most of the songs come from the 90s, only one from the 70s, which one is it?

```{r}
rankings %>% filter(decade == 1970)
```

What are the best years in terms of rating?

```{r}
rankings %>% 
  group_by(year) %>% 
  summarise(total_points = sum(points)) %>% 
  ggplot(aes(x=year,y=total_points)) +
  geom_point() +
  geom_line()
```

The ratings spiked around the nineteen-nineties and have been on a decline since.

```{r}
rankings %>% 
  group_by(decade) %>% 
  summarise(max = max(points))
```

There is however a song in the last decade that surpassed the best rated one in the 
decade before.

```{r}
rankings %>% 
  group_by(decade) %>% 
  do(arrange(.,desc(points))) %>% 
  slice(1)
```

Runaway by Kanye West ft. Pusha T is rated higher than Rapper's Delight and B.O.B.

Which artist has the most titles in the dataset?

```{r}
rankings %>% 
  count(artist,sort = TRUE)
```

Jay-Z has the most rated titles - 15.

The majority of artists has only one title rated.

```{r}
rankings %>% 
  count(artist) %>% 
  count(n)
```

```{r}
rankings %>% 
  group_by(artist) %>% 
  summarise(n_decades = n_distinct(decade)) %>% 
  arrange(desc(n_decades))
```

Jay-Z is also the only one who has a title rated within three decades period.

We know that The Notorious B.I.G. has the best rated title, does it also mean he
has the most points in total?

```{r}
rankings %>% 
  group_by(artist) %>% 
  summarise(total_points = sum(points)) %>% 
  arrange(desc(total_points)) %>% 
  top_n(15)
```

Yes! The Notorious B.I.G. has the most points in total as well as the best rated title.

It still holds in terms of average rating.

```{r}
rankings %>% 
  group_by(artist) %>% 
  summarise(mean_points = mean(points), n = n()) %>% 
  arrange(desc(mean_points)) %>% 
  top_n(15)
```

 Is there any title ranked in more then one year?
 
```{r}
rankings %>% 
  count(title, sort = TRUE)
```
 
 There are two of them:
 
```{r}
rankings %>% filter(title %in% c('All Of The Lights','Hip Hop'))
```
 
 So Hip Hop is a name shared by two distinct titles, however it looks the All Of The Lights has a male as well as mixed version, both from 2010.
 
```{r}
rankings %>% 
  mutate(n_ratings = n1 + n2 + n3 + n4 + n5) %>% 
  group_by()
```

## Gender related questions

What is the ratio of men-women?

```{r}
rankings %>% 
  count(gender)
```

Whats the gender average rating?

```{r}
rankings %>% 
  group_by(gender) %>% 
  summarise(avg_rating = mean(points))
```

Are there years when female artists had the best rated title?

```{r}
rankings %>% 
  group_by(year) %>% 
  do(arrange(.,desc(points))) %>% 
  slice(1) %>% 
  filter(gender=="female")
```

There are five years when female artists had the best rated title.

```{r}
rankings %>% 
  filter(gender=="female") %>% 
  arrange(desc(points))
```

The best-rated female title however was not the best rated title in the year 1993.

What about a year when female artists would have the most points in total or
the better average rating?

```{r}
rankings %>% 
  group_by(gender, year) %>% 
  summarise(sum_points = sum(points), avg_points = mean(points)) %>% 
  ungroup() %>% 
  group_by(year) %>% 
  mutate(rank_sum = rank(desc(sum_points)),
         rank_avg = rank(desc(avg_points))) %>% 
  filter(gender=="female" & (rank_sum==1|rank_avg==1))
```

1980 is the only year when female artists had the most points in total. There are 8 years when female titles had higher avgerage rating then male or mixed artists.

Which female artist has :
 - the most titles
 - the most points 
 - the best avg rating
 
```{r}
rankings %>% 
  filter(gender=="female") %>% 
  group_by(artist) %>% 
  summarise(n_songs=n(), points_total=sum(points), points_avg=mean(points)) %>% 
  arrange(desc(n_songs))
```
 
 Missy Elliott has the most songs, but Lauryn Hill has the best rated one.
 
 What is the first year a female title is rated?
 
```{r}
rankings %>% 
  filter(gender=="female") %>% 
  arrange(year) %>% 
  top_n(10)
```
 
 
 